Network Configuration Synthesis with Abstract Topologies

### 2016-11-15 paper/abstraction/paper.tex
more edits

\newcommand{\sysname}{{\small \sf Methane}\xspace}
\newcommand{\sysnamesec}{{\sf Methane}\xspace}
\newcommand{\propane}{{\small \sf Propane}\xspace}
\newcommand{\mbgp}{{\small \sf mBGP}\xspace}

@@

\newcommand{\CD}[1]{\texttt{\small #1}}  % code font
\newcommand{\KW}[1]{\texttt{\small\bfseries{#1}}}

@@

\textbf{Abstract---}
We develop \sysname, a system to generate provably-correct configurations for large, evolving networks from high-level specifications of their topology, routing policy, and fault-tolerance requirements. It is based on new abstractions for capturing parameterized network topologies and their evolution, and algorithms to analyze the combined impact of topology and routing policy on the network's fault tolerance. Our analysis operates entirely on abstract topologies and guarantees correctness for all concrete instantiations of the topology. Its output is also guaranteed to require minimal changes to the configurations of existing devices when the topology evolves to include new devices or links. Our experiments with many real-world topologies and policies show that our abstractions and fault-tolerance analysis are effective and that, for large networks, configuration synthesis with \sysname is two orders of magnitude faster than similar systems that operate over concrete topologies.

@@

Based on these inputs, the \sysname compiler generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the resulting
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how fault-tolerance analysis can be done in the abstract domain by using a set of sound inference rules to infer the lower bound on the number of edge-disjoint paths in any concrete topology.
%Our analysis uses a set of sound inference rules and  using an optimizing SMT solver.

@@

Second, \sysname can scale substantially better than its nearest competitor \propane that operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgery}.
Consequently, the \sysname compiler can be two orders of magnitude
faster than \propane, taking less than 10 seconds to synthesize templates for large structured networks.

%\todo{Talk about evaluation result here ... roll the point about faster compilation there.}

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, our contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item New topology abstractions for network programming
based on graph homomorphisms and connectivity rules.
\item New algorithms for analyzing routing policy and fault tolerance over
abstract topologies.
%one which guarantees correct compilation and one which determines fault tolerance
%properties of network policy.
%Both algorithms provide guarantees for any
%concrete network satisfying the abstract connectivity invariants.
%\item Design of a template generation strategy that supports network
%evolution.
\item Implementation and evaluation of the \sysname compiler that generates
provably-correct templates and is much faster
than a compiler based on concrete topologies.
\end{itemize}

@@

In this work, we focus on synthesizing configurations for a network that runs the Border Gateway Protocol (BGP), the standard for routing across independent organizations and inside large datacenters. It scales effectively and allows operators to define flexible policies.

@@

All route announcements carry an {\em AS path}, the sequence of BGP speakers (called Autonomous Systems) that will be traversed to reach the destination.
%When BGP is used inside a datacenter, each router tends to be its own AS.
Optionally, they may also include one or more {\em communities}, which are arbitrary tags whose semantics are agreed upon out of band.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Idealized small configuration component for the datacenter spine role based on router templates from a large cloud provider (\IE, macros over gobbledygook).}
  \label{fig:templates}
  \vspace{-.8em}
\end{figure}

@@

To simplify network configuration, operators are adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration per router, operators author a template per role.
%
A {\em role} is a specific function that is served by one or more routers.
%\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor configuration languages.}
%
For example, the network in Figure~\ref{fig:example} might have five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Figure~\ref{fig:templates} shows an example of what a small component of a template for the spine role in the two datacenters might look like. The template has parameters for various aspects of the configuration (\EG, neighbor list, local prefixes) and are compiled to low-level device configurations by instantiating the parameters using the network topology and a database of network information.

@@

%While templates help avoid a class of configuration errors (e.g., inconsistent routing adjacencies), they provide no guarantees
As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. Worse, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

%Ideally, we would like to find a suitable topology abstraction that can describe both of the datacenters from figures~\ref{fig:example} and~\ref{fig:example2} as well as any reasonable generalization of these datacenters. \sysname provides several abstraction mechanisms to achieve this goal.

@@

Our topology abstractions can trivially capture concrete topologies; there will be a one-to-one correspondence between abstract and concrete nodes (and edges). This ability enables operators to define complex networks in which they expect some (e.g., legacy) parts of the network to not evolve and others to evolve. It also means that our compiler subsumes those that operate only over concrete topologies.

@@

%Although the concrete \sysname policy can help prevent a number of configuration errors,
%the policy is intricately linked to the actual structure of the concrete network.
%For example, each prefix in the routing constraint is enumerated with its concrete destination
%location, and the definition of the valleyfree constraint depends on the actual routers that
%comprise each tier of the datacenter.
%
%This is unfortunate since any expansion of our datacenter network will require a corresponding
%change to the policy. Furthermore, there is no guarantee that a safe policy can be found
%for the expanded network, and even if one can be found, it may require changes to every
%configuration in the network -- a huge barrier to adoption in practice.

@@

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. It says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. As we will see in \S\ref{sec:language}, constraints like $\CD{\End(TG)}$ are just syntactic sugar for regular expressions describing network paths. The second line has a similar policy for local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the datacenter, ending at some outside location, through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

Next we can capture constraint (4) that traffic for local prefixes must stay within in the datacenter:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the variable $\CD{\$LP}$ and adds the constraint that the traffic must follow a path that matches an internal location (\IE, in the datacenter) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each datacenter level.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the datacenter through a peer.

@@

\begin{code}
Routing & Local & NoTransit &
NoValley & \Agg(PG, \In -> \Out)
\end{code}
\noindent
This instructs \sysname to satisfy the conjunction of all the constraints. It also declares that we want to perform route aggregation, as covering prefix PG, for global prefixes at the border of the datacenter (\IE, along any edge that connects the inside of the datacenter to the outside).

@@

\para{Synthesis}

%\sysname builds on the Propane network configuration tool, by supporting both synthesis for and reasoning over parameterized classes of topologies and their interactions with user-specified routing policy.

%Although \sysname can ensure correct forwarding behavior independent of the exact concrete topology or failure scenario, it is often useful to be able to ask "what if" questions about other network properties. For example, an operator may require a certain level fault tolerance between between different collections of routers. Such analysis depends, not only on the network topology, but also on the underlying routing policy, both of which \sysname has access to.
The \sysname compiler generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on the \propane compiler (for concrete networks); we show that it can be extended to abstract inputs and that the extensions are correct.

@@

Figure~\ref{fig:example-compilation} shows the PG for the datacenter policy that applies to all external traffic ($\CD{\True \Path \Exit(N1~\Prefer~N2)}$).
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\CD{\Exit(N1)}$ and the second automaton represents the less preferred constraint $\CD{\Exit(N2)}$. The PG is shown for both an instance of a simple concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstract topology.

@@

    \node[] at (#1 + 1.2, 1.25 + 2.5) {\textbf{I-local}};

    \derivation{#1}{X}{Y}
    \derivation{#1+2+#2}{L$_P$ X}{L$_P$ Y}
    %\inferencenobox{#1}{X}{}{Y}{}{}{}
    %\inferencenobox{#1+2 + #2}{$\ldots$ L$_P$ X}{}{$\ldots$ L$_P$ Y}{}{}{}
    \node[anchor=east] at (#1 + 2 + #2 - .6, 2.5) {P};
    \draw[dashed] (#1+1 + #2 - .2, -.34) rectangle (#1+3 + #2 + .1, 2.85);
}

\newcommand{\inferenceglobal}[2]{
    \node[] at (#1 + 1.2, 2.5 / 2 + 2.5) {\textbf{I-global}};
    %\inferencenobox{#1}{X}{}{Y}{}{}{}
    %\inferencenobox{#1+2+#2}{L$_P$ X}{}{A$_Q$ Y}{}{}{}
    \derivation{#1}{X}{Y}
    \derivation{#1+2+#2}{L$_P$ X}{A$_Q$ Y}

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    \inference{0}{L$_m(j,k)$}{$e_1 > 0$}{S$_n(min(j*k,e_1),1)$}{I-out1}{$e_1$}{$e_2$}
    \inference{3.5}{A$_m(j,k)$}{$e_2 > 0$}{A$_n(1,min(j,e2))$}{I-out2}{$e_1$}{$e_2$}
    \inference{7}{L$_m(j,k)$}{$e_1=n$}{A$_n(min(j*k,n),1)$}{I-mesh1}{$e_1$}{$e_2$}
    \inference{10.5}{L$_m(j,k)$}{$e_1=n$}{A$_n(1,j)$}{I-mesh2}{$e_1$}{$e_2$}
    \inferencenobox{14}{L$(j,k)$}{$mincut(X)$}{A$(1,min(k,X))$}{I-mincut}{}{}
  \end{tikzpicture}
  \end{minipage}

@@

The annotation $mincut(X)$ appearing on an edge is an assertion about the fault tolerance between nodes in two different roles. The rule \infrule{I-mincut} uses such assertions. To each node in role $n$, from a node in role $m$, we can construct at least the minimum of $X$ and $k$ disjoint paths.

@@

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is useful particularly useful for datacenter topologies where routers in one tier of the datacenter often have a particular striping pattern with another tier.

@@

Recall the policy for global prefixes in the datacenter.
%
\begin{code}
\$GP \Path \End(TG) \ensuremath{\cap}
       \Novalley(\{TG,TL\},\{AG,AL\},\{S\}) \ensuremath{\cap}
       !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent
%
Figure~\ref{fig:example-inference} shows part of the abstract PG representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially).
%
The first step applies each inference rule to this initial fact. Rules \infrule{I-mesh1} together with \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles results in a call to $\nu$Z to minimize the expression $min(\infty, AG)$, which results in 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The algorithm will then eventually apply \infrule{I-out2} to learn that any single spine node is reachable at $(S,0)$. It will also apply \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at $(S,0)$ and that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.

@@

The algorithm was able to infer that there is at least 1 disjoint path to any spine router, and at least 2 disjoint paths to any TL router. In this case, the analysis is precise. There exists a concrete network, namely the datacenter from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint.
%The bound on the fault tolerance between ToR and spine routers can also be used to determine that an aggregation-induced black hole may arise at the datacenter boundary after a single failure in some concrete instantiation.

@@

Suppose an operator wants to expand the concrete datacenter by adding an additional ToR router to the TG role. Per the network routing policy, the new ToR router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine router configuration.

@@

The \sysname compiler is implemented in around 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate configurations for open-source Quagga~\cite{quagga} routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to test validity and minimize variables subject to the topology constraints. Since the analysis typically calls the SMT solver many times with relatively small optimization problems, we use a timeout of 200ms.

@@

We evaluate \sysname by applying it on policies for backbone and datacenter networks. We aim to evaluate the expressiveness of our abstractions, the precision of the abstract disjoint-path analysis, and the generation time of \sysname both with abstraction and without.

@@

%\para{Fat Tree Topologies}
%
%The fat tree topology is a widely used datacenter topology. We look at the precision of our analysis on four fat tree variants used in practice: a standard fat tree~\cite{fattree}, the Facebook fat tree~\cite{facebook-fattree}, the F10 fault tolerant fat tree~\cite{f10-fattree}, and the VL2 topology~\cite{vl2-fattree}. For each fat tree variant, we use a tiered abstraction similar to that in our previous example. Each fat tree topology is parameterized over the number of pods $k$, which can be scaled up in the abstraction to allow for expansion.

@@

\para{Recursive Topologies}
These topologies include BCube~\cite{bcube} and DCell~\cite{dcell}. Each topology includes a recursion depth parameter ($k$), which we fixed while abstracting them. For a recursive topology with depth $k$, we model it as an abstract topology consisting of a pod to represent all depth $k-1$ subcomponents. This allows for safe expansion within a subcomponent, but does not allow changing the recursion depth dynamically.
For BCube, we model each tier of the datacenter as a separate role. Figure~\ref{fig:example-abstractions} shows an example of a BCube abstraction for $k=1$.

@@

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-time.png}} \\
  \vspace{-.6em}
  \caption{Concrete vs. Abstract Synthesis Time. \label{fig:compilation-times}}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-analysis-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-analysis-time.png}} \\
  \vspace{-.6em}
  \caption{Abstract Synthesis Time by Phase. \label{fig:abstract-breakdown}}
  \vspace{-1em}
\end{figure}

We evaluate synthesis time in \sysname using routing policy for backbone and datacenter networks inspired by configurations from a large cloud provider. For both types of networks, we fix the routing policy and scale the topology.

\para{Datacenter}

Routers in the datacenter network run BGP using unique AS numbers and connect to multiple external neighbors. The routers aggregate some prefix blocks when announcing them to external neighbors, and keep some prefixes internal. The data center prefers that traffic leave through certain neighbors over others and should not transit traffic between neighbors. The policy also says that the datacenter routers not use external neighbors to reach ``private" destinations  (i.e., those in the part of the IP address space reserved for private use).

We use a fat tree~\cite{fattree} and scale it by increasing the number of pods. The abstract topology uses one abstract node for each tier with additional nodes for local and global ToRs.

>>>>

\newcommand{\sysname}{{\text{}\small \sf Methane}\xspace}
\newcommand{\sysnamesec}{{\sf Methane}\xspace}
\newcommand{\mbgp}{{\small \sf mBGP}\xspace}

@@

\newcommand{\CD}[1]{\texttt{\small #1}}
\newcommand{\KW}[1]{\texttt{\small\bfseries{#1}}}

@@

\textbf{Abstract---}
We develop \sysname, a system to generate provably-correct configurations for large, evolving networks from high-level specifications of their topology, routing policy, and fault-tolerance requirements. It is based on new abstractions for capturing parameterized network topologies and their evolution, and algorithms to analyze the combined impact of topology and routing policy on the network's fault tolerance. Our analysis operates entirely on abstract topologies and guarantees correctness for all concrete instantiations of the topology. Its output is also guaranteed to require minimal changes to the configurations of existing devices when the topology evolves to include new devices or links. Our experiments with many real-world topologies and policies show that our abstractions and fault-tolerance analysis are effective and that, for large networks, configuration synthesis with \sysname is two orders of magnitude faster than systems that operate over concrete topologies.

@@

Based on these inputs, \sysname generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the resulting
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how fault-tolerance analysis can be done in the abstract domain by using a set of sound inference rules to infer the lower bound on the number of edge-disjoint paths in any concrete topology.
%Our analysis uses a set of sound inference rules and  using an optimizing SMT solver.

@@

Second, \sysname can scale substantially better than its nearest competitor Propane that operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgery}.
Consequently, \sysname can be two orders of magnitude
faster than Propane, taking less than 10 seconds to synthesize templates for large structured networks.

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, our contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item New topology abstractions for network programming
based on graph homomorphisms and connectivity rules.
\item New algorithms for analyzing routing policy and fault tolerance over
abstract topologies.
%one which guarantees correct compilation and one which determines fault tolerance
%properties of network policy.
%Both algorithms provide guarantees for any
%concrete network satisfying the abstract connectivity invariants.
%\item Design of a template generation strategy that supports network
%evolution.
\item Implementation and evaluation of the \sysname tool that generates
provably-correct templates and is much faster
than tools based on concrete topologies.
\end{itemize}

@@

In this work, we focus on synthesizing configurations for a network that runs the Border Gateway Protocol (BGP), the standard for routing across independent organizations and inside large data centers. It scales effectively and allows operators to define flexible policies.

@@

All route announcements carry an {\em AS path}, the sequence of BGP speakers (called Autonomous Systems) that will be traversed to reach the destination.
%When BGP is used inside a data center, each router tends to be its own AS.
Optionally, they may also include one or more {\em communities}, which are arbitrary tags whose semantics are agreed upon out of band.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/templates}
  \caption{Idealized small configuration component for the data center spine role based on router templates from a large cloud provider (\IE, macros over gobbledygook).}
  \label{fig:templates}
  \vspace{-.8em}
\end{figure}

@@

To simplify network configuration, operators are adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration per router, operators author a template per role.
%
A {\em role} is a specific function that is served by one or more routers.
%\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor configuration languages.}
%
For example, the network in Figure~\ref{fig:example} might have five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Figure~\ref{fig:templates} shows an example of what a small component of a template for the spine role in the two data centers might look like. The template has parameters for various aspects of the configuration (\EG, neighbor list, local prefixes) and are compiled to low-level device configurations by instantiating the parameters using the network topology and a database of network information.

@@

As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. Worse, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

%Ideally, we would like to find a suitable topology abstraction that can describe both of the data centers from figures~\ref{fig:example} and~\ref{fig:example2} as well as any reasonable generalization of these data centers. \sysname provides several abstraction mechanisms to achieve this goal.

@@

Our topology abstractions can also capture concrete topologies; there will be a one-to-one correspondence between abstract and concrete nodes (and edges). This ability enables operators to define complex networks in which they expect some (e.g., legacy) parts of the network to not evolve and others to evolve. 
%It also means that our compiler subsumes those that operate only over concrete topologies.

@@

%Although the concrete \sysname policy can help prevent a number of configuration errors,
%the policy is intricately linked to the actual structure of the concrete network.
%For example, each prefix in the routing constraint is enumerated with its concrete destination
%location, and the definition of the valleyfree constraint depends on the actual routers that
%comprise each tier of the data center.
%
%This is unfortunate since any expansion of our data center network will require a corresponding
%change to the policy. Furthermore, there is no guarantee that a safe policy can be found
%for the expanded network, and even if one can be found, it may require changes to every
%configuration in the network -- a huge barrier to adoption in practice.

@@

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. It says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. As we will see in \S\ref{sec:language}, constraints like $\CD{\End(TG)}$ are just syntactic sugar for regular expressions describing network paths. The second line has a similar policy for local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the data center, ending at some outside location, through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

Next we can capture constraint (4) that traffic for local prefixes must stay within in the data center:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the variable $\CD{\$LP}$ and adds the constraint that the traffic must follow a path that matches an internal location (\IE, in the data center) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each data center level.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the data center through a peer.

@@

\begin{code}
Routing & Local & NoTransit &
NoValley & \Agg(PG, \In -> \Out)
\end{code}
\noindent
This instructs \sysname to satisfy the conjunction of all the constraints. It also declares that we want to perform route aggregation, as covering prefix PG, for global prefixes at the border of the data center (\IE, along any edge that connects the inside of the data center to the outside).

@@

% The \text{} is needed for some reason for formatting
\para{Synthesis} \sysname generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on Propane (for concrete networks); we show that it can be extended to abstract inputs and that the extensions are correct.

@@

Figure~\ref{fig:example-compilation} shows the PG for the data center policy that applies to all external traffic ($\CD{\True \Path \Exit(N1~\Prefer~N2)}$).
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\CD{\Exit(N1)}$ and the second automaton represents the less preferred constraint $\CD{\Exit(N2)}$. The PG is shown for both an instance of a simple concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstract topology.

@@

    \node[] at (#1 + 1.2, 1.25 + 2.5) {\textbf{I-local}};
    \derivation{#1}{X}{Y}
    \derivation{#1+2+#2}{L$_P$ X}{L$_P$ Y}
    \node[anchor=east] at (#1 + .5 + #2, 2.5 / 2) {$\Rightarrow$};
    %\inferencenobox{#1}{X}{}{Y}{}{}{}
    %\inferencenobox{#1+2 + #2}{$\ldots$ L$_P$ X}{}{$\ldots$ L$_P$ Y}{}{}{}
    \node[anchor=east] at (#1 + 2 + #2 - .6, 2.5) {P};
    \draw[dashed] (#1+1 + #2 - .2, -.34) rectangle (#1+3 + #2 + .1, 2.85);
}

\newcommand{\inferenceglobal}[2]{
    \node[] at (#1 + 1.2, 2.5 / 2 + 2.5) {\textbf{I-global}};
    %\inferencenobox{#1}{X}{}{Y}{}{}{}
    %\inferencenobox{#1+2+#2}{L$_P$ X}{}{A$_Q$ Y}{}{}{}
    \derivation{#1}{X}{Y}
    \derivation{#1+2+#2}{L$_P$ X}{A$_Q$ Y}
    \node[anchor=east] at (#1 + .5 + #2, 2.5 / 2) {$\Rightarrow$};

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    \inference{0}{L$_m(j,k)$}{$e_1 > 0$}{S$_n(min(j*k,e_1),1)$}{I-out1}{$e_1$}{$e_2$}
    \inference{3.5}{A$_m(j,k)$}{$e_2 > 0$}{A$_n(1,min(j,e2))$}{I-out2}{$e_1$}{$e_2$}
    \inference{7}{L$_m(j,k)$}{$e_1=n$}{A$_n(min(j*k,n),1)$}{I-mesh1}{$e_1$}{$e_2$}
    \inference{10.5}{L$_m(j,k)$}{$e_1=n$}{A$_n(1,j)$}{I-mesh2}{$e_1$}{$e_2$}
    \inferencenobox{14}{L$(j,k)$}{$\Mincut(X)$}{A$(1,min(k,X))$}{I-mincut}{}{}
  \end{tikzpicture}
  \end{minipage}

@@

The annotation $\Mincut(X)$ appearing on an edge is an assertion about the fault tolerance between nodes in two different roles. The rule \infrule{I-mincut} uses such assertions. To each node in role $n$, from a node in role $m$, we can construct at least the minimum of $X$ and $k$ disjoint paths.

@@

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is useful particularly useful for data center topologies where routers in one tier of the data center often have a particular striping pattern with another tier.

@@

Recall the policy for global prefixes in the data center.
%
\begin{code}
\$GP \Path \End(TG) \ensuremath{\cap}
       \Novalley(\{TG,TL\},\{AG,AL\},\{S\}) \ensuremath{\cap}
       !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent
%
Figure~\ref{fig:example-inference} shows part of the abstract PG representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially).
%
The first step applies each inference rule to this initial fact. Rules \infrule{I-mesh1} together with \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles results in a call to $\nu$Z to minimize the expression $min(\infty, AG)$, which results in 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The algorithm will then eventually apply \infrule{I-out2} to learn that any single spine node is reachable at $(S,0)$. It will also apply \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at $(S,0)$ and that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.

@@

The algorithm was able to infer that there is at least 1 disjoint path to any spine router, and at least 2 disjoint paths to any TL router. In this case, the analysis is precise. There exists a concrete network, namely the data center from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint. 
%The bound on the fault tolerance between ToR and spine routers can also be used to determine that an aggregation-induced black hole may arise at the data center boundary after a single failure in some concrete instantiation.

@@

Suppose an operator wants to expand the concrete data center by adding an additional ToR router to the TG role. Per the network routing policy, the new ToR router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine router configuration.

@@

\sysname is implemented in approximately 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate configurations for open-source Quagga~\cite{quagga} routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to test validity and minimize variables subject to the topology constraints. Since the analysis typically calls the SMT solver many times with relatively small optimization problems, we use a timeout of 200ms.

@@

We evaluate \sysname by applying it on policies for backbone and data center networks. We aim to evaluate the expressiveness of our abstractions, the precision of the abstract disjoint-path analysis, and the generation time of \sysname both with abstraction and without.

@@

%\para{Fat Tree Topologies}
%
%The fat tree topology is a widely used data center topology. We look at the precision of our analysis on four fat tree variants used in practice: a standard fat tree~\cite{fattree}, the Facebook fat tree~\cite{facebook-fattree}, the F10 fault tolerant fat tree~\cite{f10-fattree}, and the VL2 topology~\cite{vl2-fattree}. For each fat tree variant, we use a tiered abstraction similar to that in our previous example. Each fat tree topology is parameterized over the number of pods $k$, which can be scaled up in the abstraction to allow for expansion.

@@

\para{Recursive Topologies}
These topologies include BCube~\cite{bcube} and DCell~\cite{dcell}. Each topology includes a recursion depth parameter ($k$), which we fixed while abstracting them. For a recursive topology with depth $k$, we model it as an abstract topology consisting of a pod to represent all depth $k-1$ subcomponents. This allows for safe expansion within a subcomponent, but does not allow changing the recursion depth dynamically.
For BCube, we model each tier of the data center as a separate role. Figure~\ref{fig:example-abstractions} shows an example of a BCube abstraction for $k=1$.

@@

\begin{figure}[t!]
  \subcaptionbox{Data center}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-time.png}} \\
  \vspace{-.6em}
  \caption{Concrete vs. Abstract Synthesis Time. \label{fig:compilation-times}}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \subcaptionbox{Data center}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-analysis-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-analysis-time.png}} \\
  \vspace{-.6em}
  \caption{Abstract Synthesis Time by Phase. \label{fig:abstract-breakdown}}
  \vspace{-1em}
\end{figure}

We evaluate generation time in \sysname both with and without abstraction using routing policy for backbone and data center networks inspired by configurations obtained from a large cloud provider. For both types networks, we fix the routing policy and scale the size of the underlying topology.

\para{Data center}

Routers in the data center network run BGP using private AS numbers and peer with multiple backbone networks. The routers aggregate some prefix blocks when announcing them to the backbone network, and keep some prefixes internal. The policy also ensures that private address space from external neighbors is dropped. The data center prefers that traffic leave through certain peers over others and ensures that transit traffic between peers is dropped.

We use a fat tree topology and scale the size of the topology according the the number of pods $k$. The abstract topology uses one abstract node for each tier of the data center with an additional node for separate local and global ToRs.
